Temporality of the Victorian Hospital Patient Transfer Network

### 2025-05-27 reports/paper.qmd
Update the home prop fig

```{python}
#| label: compute-imports
import numpy as np
from pathlib import Path
import polars as pl
from matplotlib import pyplot as plt, dates as mdates
import seaborn as sns
import calendar
from itertools import accumulate
import datetime

@@

axs[-1].set_xlabel('Extent')
```

>>>>

```{python}
#| label: compute-imports
import numpy as np
from pathlib import Path
import polars as pl
from matplotlib import pyplot as plt, dates as mdates
import seaborn as sns
import calendar
from itertools import accumulate
from collections import defaultdict
import datetime

@@

_ = axs[-1].set_xlabel('Extent')
```

One striking feature

```{python}
histories = {
  dir.stem: np.loadtxt(dir/"hosp_presence.csv", delimiter=',')
  for dir in (root / "simulations/zero_sims_resized/" ).iterdir()
}

_BASE_DATE = np.datetime64('2011-01', 'M')
_MAX_DATE = np.datetime64('2018-12', 'M')
def set_xdatelabels(hist, ax, spacing=12, dt=1.0):
    # here, we want to get succinct date formats
    # we map the times onto the timedeltas of the simulations
    N = hist.shape[1]
    emp_max_date = (np.timedelta64(int(N*dt), 'D') + _BASE_DATE).astype('datetime64[M]')
    ax_max_date = np.min([emp_max_date, _MAX_DATE])

    dates = np.arange(_BASE_DATE, ax_max_date, np.timedelta64(spacing, 'M'))

    label_pos = (dates.astype('datetime64[D]').astype('int64') - _BASE_DATE.astype('datetime64[D]').astype('int64')) / dt

    ax.set_xticks(label_pos, dates.astype('datetime64[Y]'), rotation=45, rotation_mode='anchor')

    return ax.get_xticks()

def heatmap(hist, nmax=30):
    # hist is 2D
    bins = np.arange(nmax+2)
    hist_heat = np.array([
        np.histogram(hist[:, i], bins=bins)[0]
        for i in range(hist.shape[1])
    ]).T

    return bins, hist_heat
```

```{python}
def mega_heatmap(hists, vmax=1500):
  hms = {model: heatmap(hist_vals)[1] for model, hist_vals in hists.items()}
  fig, axs = plt.subplots(nrows=4, figsize=[16, 9], sharex=True)
  for ax, (k, hmv) in zip(axs.flatten(), hms.items()):
      if k == 'temporal':
        hmv = hmv[:,::2]
      sns.heatmap(data=hmv[::-1,:], vmax=vmax, ax=ax, cmap='inferno', cbar=False)
      ax.set_title(k)
      ax.set_ylim([-0.5, 30.5])
      ax.set_yticks([0, 10, 20, 30])
  set_xdatelabels(hmv, ax)
  fig.colorbar(ax.collections[0], ax=axs, label='Count', extend='max')

  return fig

_ = mega_heatmap(histories)
```
